---
title: "Untitled"
author: "Tashfeen Ahmed"
date: "2023-01-30"
output:
  word_document: default
  html_document: default
---
import library and datsets available in folder
```{r}
install.packages("tidyverse", repos = "http://cran.us.r-project.org")
library(tidyverse)
install.packages("dplyr", repos = "http://cran.us.r-project.org")
library(dplyr)
install.packages("janitor", repos = "http://cran.us.r-project.org")
library(janitor)
library(ggplot2)
library(lubridate)


```
Importing data

```{r}
daily_activity <- read.csv("C:\\Users\\tashf\\OneDrive\\Desktop\\Case Study\\Google_Project\\Fitabase_Data_4.12.16-5.12.16\\dailyActivity_merged.csv")
daily_sleep <- read.csv("C:\\Users\\tashf\\OneDrive\\Desktop\\Case Study\\Google_Project\\Fitabase_Data_4.12.16-5.12.16\\sleepDay_merged.csv")
weight_log <- read.csv("C:\\Users\\tashf\\OneDrive\\Desktop\\Case Study\\Google_Project\\Fitabase_Data_4.12.16-5.12.16\\weightLogInfo_merged.csv")
daily_calories <- read.csv("C:\\Users\\tashf\\OneDrive\\Desktop\\Case Study\\Google_Project\\Fitabase_Data_4.12.16-5.12.16\\dailyCalories_merged.csv")
daily_intensities <- read.csv("C:\\Users\\tashf\\OneDrive\\Desktop\\Case Study\\Google_Project\\Fitabase_Data_4.12.16-5.12.16\\dailyIntensities_merged.csv")
daily_steps <- read.csv("C:\\Users\\tashf\\OneDrive\\Desktop\\Case Study\\Google_Project\\Fitabase_Data_4.12.16-5.12.16\\dailySteps_merged.csv")

```

Data is joined using id and date columns

```{r}
merge_1 <- merge(daily_activity, daily_sleep, by = c("Id" ))
merge_2 <- merge(daily_steps, daily_calories, by = c("Id" ))
merge_3 <- merge(daily_intensities, weight_log, by = c("Id" ))
merge_4 <- merge(merge_1, merge_2, by = c("Id" ))



```

Convert Date from charater format to Date format

```{r}
merge_4$Date <- mdy(merge_4$Date)

```


summarize the data.
```{r}
summary(merge_4)

```
check data for unique id
```{r}
 n_distinct(daily_activity$Id)
nrow(daily_activity)
```
should be 30 as survey was of 30 people

```{r}
summary(merge_3)
```

quick summary statistics
```{r}
colnames(daily_activity)
```

```{r}
daily_sleep %>%  
  select(TotalSleepRecords,
  TotalMinutesAsleep,
  TotalTimeInBed) %>%
  summary()
```
```{r}
colnames(merge_3)

```

```{r}
colnames(merge_4)
```
```{r}
ggplot(data=merge_4) +geom_point(mapping=aes(x=TotalMinutesAsleep,y=TotalTimeInBed))
```
Linear trend found which is expected
```{r}
ggplot(data=merge_3) +geom_smooth(mapping=aes(y=WeightKg, x = SedentaryMinutes))
ggplot(data=merge_3) +geom_smooth(mapping=aes(x=WeightKg, y = LightlyActiveMinutes))
ggplot(data=merge_3) +geom_smooth(mapping=aes(x=WeightKg, y = FairlyActiveMinutes))
ggplot(data=merge_3) +geom_smooth(mapping=aes(x=WeightKg, y = VeryActiveMinutes))
ggplot(data=merge_3) +geom_smooth(mapping=aes(x=WeightKg, y = VeryActiveDistance))



```
From these plot we can see 
higher weight people are more sedentary. so we should target specifically above 70kg
between 90 and 120kg people are very or fairly active- seems like high weight people are trying to lose weight and exercise more than normal people but they have less very active distance which means they run/jog less and are using indoor activities  to stay active such as gym


```{r}
ggplot(data=merge_4) +geom_smooth(mapping=aes(x=WeightKg, y = StepTotal))
ggplot(data=merge_4) +geom_smooth(mapping=aes(x=WeightKg, y = TotalMinutesAsleep))
ggplot(data=merge_4) +geom_smooth(mapping=aes(x=Calories, y = StepTotal))
```

step total decreases with weight, above 100 it is very less
people who take more calories have less total 

##Conclusion target less than 70kg as they are active but wont be willing to pay a lot because they are not passionate, they have more active distance though meaning they run/walk more. however between 90 and 120kg people are passionate and would be willing to spend more money 

```{r}
heart_rate <- read.csv("heartrate_seconds_merged.csv")
hourly_calories <- read.csv("hourlyCalories_merged.csv")
hourly_steps <- read.csv("hourlySteps_merged.csv")

```
column names

```{r}
colnames(heart_rate)
colnames(hourly_calories)
colnames(hourly_steps)
```
unique ids
```{r}
n_distinct(heart_rate$Id)
n_distinct(hourly_calories$Id)
n_distinct(hourly_steps$Id)
```
Unique id should be 33

```{r}
merge_4$month <- month(merge_4$Date)
merge_4$days <- day(merge_4$Date)

```



```{r}
ggplot(data=merge_4)+ geom_bar(mapping=aes(x=days))
ggplot(data=merge_4)+ geom_bar(mapping=aes(x=month))
```


people most active in start of month and middle of month
while most active months are April and may

```{r}
ggplot(data=merge_4)+ geom_boxplot(mapping=aes(x=WeightKg, y= Time))
```
generally people data was collected at 12 mid night, 1pm noon and 9pm night



