---
title: "Bellabit - Capstone Project"
author: "Tashfeen Ahmed"
date: "2023-01-30"
output:
  word_document: default
  html_document: default
  pdf_document: default
---
import library and datasets available in folder
```{r}
library(tidyverse)
# install.packages("dplyr", repos = "http://cran.us.r-project.org")
library(dplyr)
# install.packages("janitor", repos = "http://cran.us.r-project.org")
library(janitor)
library(ggplot2)
library(lubridate)
install.packages("openxlsx", dependencies=TRUE, repos = "http://cran.us.r-project.org")


```
Importing data

```{r}
daily_activity <- read.csv('dailyActivity_merged.csv')
                           
daily_sleep <- read.csv('sleepDay_merged.csv')

weight_log <- read.csv('weightLogInfo_merged.csv')

daily_calories <- read.csv('dailyCalories_merged.csv')

daily_intensities <- read.csv('dailyIntensities_merged.csv')

daily_steps <- read.csv('dailySteps_merged.csv')

hourly_steps  <- read.csv('hourlySteps_merged.csv')

hourly_calories <- read.csv('hourlyCalories_merged.csv')

```

Preview Data

```{r}
head(daily_activity)

head(daily_sleep)


head(daily_calories)

head(weight_log)

head(hourly_steps)

head(hourly_calories)
```

```{r}
str(daily_activity)
str(daily_sleep)
str(hourly_calories)
str(weight_log)
str(hourly_steps)
str(hourly_calories)


```

To confirm a number of participants in all data sets I will count unique IDs.

```{r}
n_distinct(weight_log$Id)
         
n_distinct(daily_activity$Id)
n_distinct(daily_sleep$Id)
n_distinct(daily_calories$Id)
n_distinct(hourly_steps$Id)
n_distinct(hourly_calories$Id)
```
 checking a total number of duplicates in each data frame.
 
```{r}
sum(duplicated(weight_log))
sum(duplicated(daily_activity))
sum(duplicated(daily_sleep))
sum(duplicated(daily_calories))
sum(duplicated(hourly_steps))
sum(duplicated(hourly_calories))
```
 Only sleep data contains duplicates. They need to be cleaned.
```{r}
daily_sleep <- daily_sleep %>% 
  distinct()
sum(duplicated(daily_sleep))

```
Standardizing column names

```{r}
# clean_names(daily_activity)
# clean_names(daily_calories)
# clean_names(hourly_calories)
# clean_names(daily_sleep)
# clean_names(hourly_steps)

```

Format Date and Time

Change variable name

```{r}
daily_activity <- daily_activity %>%
    rename(ActivityDay = Date)
   
```

```{r}
daily_activity <- daily_activity %>%  
  mutate(ActivityDay = as_date(ActivityDay, format = "%m/%d/%Y"))
   
```

```{r}
daily_activity <- daily_activity %>% 
           mutate(year = lubridate::year(ActivityDay), 
                month = lubridate::month(ActivityDay), 
                day = lubridate::day(ActivityDay))


```



```{r}
daily_calories <- daily_calories %>% 
  mutate(ActivityDay = as.Date(ActivityDay, format = "%m/%d/%Y")) %>%
  mutate(year = lubridate::year(ActivityDay), 
                month = lubridate::month(ActivityDay), 
                day = lubridate::day(ActivityDay))
```

```{r}
daily_intensities <- daily_intensities %>% 
  mutate(ActivityDay = as.Date(ActivityDay, format = "%m/%d/%Y")) %>%
  mutate(year = lubridate::year(ActivityDay), 
                month = lubridate::month(ActivityDay), 
                day = lubridate::day(ActivityDay))
```

```{r}
daily_sleep <- daily_sleep %>% 
  rename(ActivityDay = SleepDay) %>%
  mutate(ActivityDay = as.Date(ActivityDay, format = "%m/%d/%Y")) %>%
  mutate(year = lubridate::year(ActivityDay), 
                month = lubridate::month(ActivityDay), 
                day = lubridate::day(ActivityDay))
```


```{r}
daily_steps <- daily_steps %>% 
  mutate(ActivityDay = as.Date(ActivityDay, format = "%m/%d/%Y")) %>%
  mutate(year = lubridate::year(ActivityDay), 
                month = lubridate::month(ActivityDay), 
                day = lubridate::day(ActivityDay))
```


```{r}
weight_log <- weight_log %>%
  rename(ActivityDay = Date) %>%
  mutate(ActivityDay = as.Date(ActivityDay, format = "%m/%d/%Y")) %>%
           mutate(year = lubridate::year(ActivityDay),
                  month = lubridate::month(ActivityDay),
                  day = lubridate::day(ActivityDay))
```


Format date/time data where a time stamp is in 12 hours AM/PM format

```{r}
daily_activity <- daily_activity %>% 
  mutate(Time = format(strptime(Time, "%H:%M"), "%I:%M"))

```




Analyze
join using id and date columns

```{r}
merge_1 <- full_join(daily_activity, daily_sleep, by = c("Id", "year", "month", "day" ))
merge_2 <- full_join(daily_steps, daily_calories, by = c("Id", "year", "month", "day" ))
merge_3 <- full_join(daily_intensities, weight_log, by = c("Id", "year", "month", "day" ))
merge_4 <- full_join(merge_1, merge_2, by = c("Id", "year", "month", "day" ))
```

```{r}
merge_5 <- full_join(merge_3,merge_4, by = c("Id", "year", "month", "day" ))
```




summarize the data.
```{r}
summary(merge_5)

```
check data for unique id and duplicate
```{r}
 n_distinct(merge_5$Id)
nrow(merge_5)
```
should be 30 as survey was of 30 people

```{r}
 # clean_names(merge_5)

```


```{r}
daily_sleep %>%  
  select(TotalSleepRecords,
  TotalMinutesAsleep,
  TotalTimeInBed) %>%
  summary()
```
```{r}
colnames(merge_5)
```

```{r}
install.packages("rio", repos = "http://cran.us.r-project.org")
library(openxlsx)
library(rio)

export(merge_5, "merge_5_to_clean.xlsx")

```



```{r}
merge_6 <- subset(merge_5, select = -c(WeightKg.y
, ActivityDay.y
, Fat.y
, LogId.y
, ActivityDay.y.x
, ActivityDay.x.y
, ActivityDay.y.y
,ActivityDay.x.x, BMI.y,WeightPounds.y,IsManualReport.y))

colnames(merge_6)
setwd("Created DataSets")
export(merge_6, "MERGE_6_CLEANED_DATA.XLSX")
```
```{r}
str(merge_6)
```




Average Table




```{r}
 average_values1 <- data.frame(avg_sedentary_mins =mean(merge_6$SedentaryMinutes),
                              
avg_light_active_mins = mean(merge_6$LightlyActiveMinutes),
                              
avg_fairly_active_mins = mean(merge_6$FairlyActiveMinutes),

avg_very_active_mins = mean(merge_6$VeryActiveMinutes),

avg_sedentary_active_distance = mean(merge_6$SedentaryActiveDistance), 
                                     
avg_light_active_distance = 3.34, 
                                                                                                          
avg_moderately_active_distance = mean(merge_6$ModeratelyActiveDistance),

avg_very_active_distance = mean(merge_6$VeryActiveDistance),

avg_minutes_sleep = mean(merge_6$TotalMinutesAsleep, na.rm = TRUE),

avg_calories = mean(merge_6$Calories),

avg_steps = mean(merge_6$StepTotal))



```
```{r}
head(average_values1)
```

```{r}
export(average_values1, "Partial Average Values.xlsx")
```


Plot Graphs



```{r}
ggplot(data=merge_6) +geom_point(mapping=aes(x=TotalMinutesAsleep,y=TotalTimeInBed))
```
Linear trend found which is expected
```{r}
ggplot(data=merge_6) +geom_smooth(mapping=aes(y=WeightKg.x, x = SedentaryMinutes))
ggplot(data=merge_6) +geom_smooth(mapping=aes(x=WeightKg.x, y = LightlyActiveMinutes))
ggplot(data=merge_6) +geom_smooth(mapping=aes(x=WeightKg.x, y = FairlyActiveMinutes))
ggplot(data=merge_6) +geom_smooth(mapping=aes(x=WeightKg.x, y = VeryActiveMinutes))
ggplot(data=merge_6) +geom_smooth(mapping=aes(x=WeightKg.x, y = LightActiveDistance))
ggplot(data=merge_6) +geom_smooth(mapping=aes(x=WeightKg.x, y = ModeratelyActiveDistance))
ggplot(data=merge_6) +geom_smooth(mapping=aes(x=WeightKg.x, y = VeryActiveDistance))


```
From these plot we can see 
higher weight people are more sedentary. so we should target specifically below 65kg
between 90 and 120kg people are very or fairly active- seems like high weight people are trying to lose weight and exercise more than normal people and they have more very active distance which means they run/jog less and are using indoor activities  to stay active such as gym


```{r}
ggplot(data=merge_6) +geom_smooth(mapping=aes(x=WeightKg.x, y = StepTotal))
ggplot(data=merge_6) +geom_smooth(mapping=aes(x=WeightKg.x, y = TotalMinutesAsleep))
ggplot(data=merge_6) +geom_smooth(mapping=aes(x=Calories, y = StepTotal))
```

step total decreases with weight, above 100kg steps decline
people who take more calories have more steps there is a linear relationship 

##Conclusion target less than 65kg as they are active but wont be willing to pay a lot because they are not passionate, they have more active distance though meaning they run/walk more. however between 90 and 120kg people are passionate and would be willing to spend more money 




```{r}
ggplot(data=merge_6)+ geom_bar(mapping=aes(x=day))
ggplot(data=merge_6)+ geom_bar(mapping=aes(x=month))
```


people most active in start of month and middle of month
while data was collected only for April and May Months

```{r}
ggplot(data=merge_6) +geom_smooth(mapping=aes(x=TotalMinutesAsleep
, y = LightlyActiveMinutes))

ggplot(data=merge_6) +geom_smooth(mapping=aes(x=TotalMinutesAsleep
, y = FairlyActiveMinutes
))

ggplot(data=merge_6) +geom_smooth(mapping=aes(x=TotalMinutesAsleep
, y = VeryActiveMinutes
))
```


Active people sleep less than 400 minutes


```{r}
heart_rate <- read.csv("heartrate_seconds_merged.csv")
hourly_calories <- read.csv("hourlyCalories_merged.csv")
hourly_steps <- read.csv("hourlySteps_merged.csv")

```
column names

```{r}
colnames(heart_rate)
colnames(hourly_calories)
colnames(hourly_steps)
```
unique ids

```{r}
n_distinct(heart_rate$Id)
n_distinct(hourly_calories$Id)
n_distinct(hourly_steps$Id)
```

Unique id should be 30 as survey was for 40 people

Format Date and Time

```{r}
heart_rate$Time <- dmy_hms(heart_rate$Time)

```
```{r}
heart_rate <- heart_rate %>%
  mutate(ActivityDay = as.Date(Time,  format = "%m/%d/%Y"))
```


```{r}
heart_rate$Time <- format(as.POSIXct(heart_rate$Time), format = "%H:%M:%S")

heart_rate <- heart_rate %>%
  mutate( year = lubridate::year(ActivityDay),
          month = lubridate::month(ActivityDay),
          day = lubridate::day(ActivityDay))
```


```{r}
hourly_calories$ActivityHour <- dmy_hms(hourly_calories$ActivityHour)

hourly_calories <- hourly_calories %>%
  mutate(ActivityDay = as.Date(ActivityHour,  format = "%m/%d/%Y"))

hourly_calories$ActivityHour <- format(as.POSIXct(hourly_calories$ActivityHour), format = "%H:%M:%S")

hourly_calories <- hourly_calories %>% 
  rename(Time = ActivityHour) %>%
  mutate( year = lubridate::year(ActivityDay),
          month = lubridate::month(ActivityDay),
          day = lubridate::day(ActivityDay))
```


```{r}
hourly_steps$ActivityHour <- dmy_hms(hourly_steps$ActivityHour)

hourly_steps <- hourly_steps %>%
  mutate(ActivityDay = as.Date(ActivityHour,  format = "%m/%d/%Y"))

hourly_steps$ActivityHour <- format(as.POSIXct(hourly_steps$ActivityHour), format = "%H:%M:%S")

hourly_steps <- hourly_steps %>% 
  rename(Time = ActivityHour)%>%
  mutate( year = lubridate::year(ActivityDay),
          month = lubridate::month(ActivityDay),
          day = lubridate::day(ActivityDay))
```
Plot Graphs


```{r}
average_values_2 <- data.frame(avg_heart_rate =mean(heart_rate$Value),
                              
avg_hourly_steps = mean(hourly_steps$StepTotal),
                              
avg_hourly_calories = mean(hourly_calories$Calories))
```


```{r}
average_values <- merge(average_values1,average_values_2)

head(average_values)
setwd("Created DataSets")
export(average_values, "Complete Average Values.xlsx")
```
merge data

```{r}
merge_7 <- merge(daily_activity, heart_rate, by = c("Id"))
```

```{r}
merge_8 <- merge(daily_activity, hourly_calories, by = c("Id"))
merge_9 <- merge(daily_activity, hourly_steps, by = c("Id"))

```

Plot Graphs


```{r}
ggplot(data=merge_9) +geom_smooth(mapping=aes(x=WeightKg, y = StepTotal))
```

```{r}
ggplot(data=merge_8) +geom_smooth(mapping=aes(x=WeightKg, y = Calories))
```

People between 80 and 100kg take most calories
